---
title: "Alpha_power_MLM"
author: "Hannah Coyle"
date: "18/10/2019"
output: html_document
---
 ##What is the effect of timepoint and group of on alpha power?
```{r global options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figures/',
                      echo=FALSE, warning=FALSE, message=FALSE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
```

```{r load libaries, include = FALSE}

library(tidyverse)
library(knitr)
library(tidyverse)
library(Rmisc)
library(lme4)
library(sjPlot)
library(dotwhisker)
library(broom)
library(car)
library(e1071)
library(ggeffects)
library(magrittr)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(lme4)
```

# Load data frame with mean alpha power from ROI electrodes across BL, T1, T2 for controls and mTBI
```{r set up file, include= FALSE}
here<-"~/Documents/PHD-Data-Analysis/PHD-Data-Analysis"
setwd(here) #main path/homedirectory
setwd("./Analysis/Data")
load("power_df_all.Rdata")
```

# Visualise data as bar and line graphs
```{r Visualise alpha power longitudinally}

alpha_df<-power_df_all %>%
  mutate(code= as.numeric(code)) %>%
  group_by(group, timepoint) 

alpha_summary<- alpha_df %>%
  group_by(group, timepoint) %>%
  dplyr::select(alpha) %>%
  set_colnames(c("group","timepoint", "alpha")) %>%
  gather(key="measure", value= "alpha_power", -group, -timepoint, na.rm=TRUE) %>%
  summarySE(measurevar="alpha_power", groupvars=c("group","measure", "timepoint"))

# quick graph as a bar graph
alpha_bar <- alpha_summary %>%
  mutate(measure=factor(measure, labels ="alpha")) %>%
  ggplot(aes(x=measure, y=alpha_power, fill=group)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=alpha_power-se, ymax=alpha_power+se),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9))+
  ylab("Alpha Power") +
  facet_grid(~timepoint)+
  theme_light() +
  scale_fill_manual(values=c("deepskyblue1","royalblue4"))+
  ggtitle(" ROI Mean Absolute Power Eyes Closed") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title.x=element_blank(),
        axis.text.x=element_blank())

print(alpha_bar)

 # quick graph as a line graph
  
  gbase<- alpha_summary %>%
    ggplot(aes(y= alpha_power, x= timepoint, colour=group))+
    geom_point()  #+ facet_grid(~group)
    #geom_errorbar(aes(ymin = alpha_power - se, ymax = alpha_power + se),
                  #width=.2) ## optional SE bars, make graph look messy though
  
  
  gline <- gbase + geom_line() 
  alpha_summary$time = as.numeric(alpha_summary$timepoint)
  gline <- gline %+% alpha_summary
  
  print(gline + aes(x=time)+
          scale_x_continuous(breaks=c(1:3), labels=c("BL", "T1", "T2")))
  
```
 Bar graph and line graphs show distribution of means  
 
```{r Visualise bar plots}
## plot individual points with means as a bar plot
base_plot <- ggplot(power_df_all, aes(x = timepoint, y = alpha)) +
  facet_grid(~group) +
  stat_summary(aes(fill = factor(timepoint)), fun.y = mean, geom = "bar", color = "black", na.rm = TRUE) + 
  scale_fill_manual(values = c("#66c2a5", "#8da0cb", "#7f66c2")) +  # colorbrewer2.org
  theme(legend.position = "none")
# put individual lines on bar graph
subj_means_plot <- base_plot + 
  stat_summary(aes(group = code), fun.y = mean,  # means from raw data
               geom = "point", shape = 19, size = 4, color = "skyblue1") +
  stat_summary(aes(group = code), fun.y = mean, 
               geom = "line", size = 1.2, color = "skyblue1")

print(subj_means_plot)
```
 Bar graph with individual points shows individual trajectories (and what are likely to be outliers)

# Check normality of outcome measure and transform  if necessary
```{r check normality of outcome measure, include= TRUE, echo= TRUE}
hist(power_df_all$alpha) # very left skewed
#Try log transforming the outcome variable
power_df_all$alpha_log <-log(power_df_all$alpha)

hist(power_df_all$alpha_log) # yes is now normally distributed

alpha_diff<- power_df_all %>%
  select(group,code,  alpha, timepoint) %>%
  spread(key=timepoint, value= alpha) %>%
  mutate(alpha_bl = 0) %>%
  mutate(alpha_t1= T1-BL) %>%
  mutate(alpha_t2= T2-BL) %>%
  select( group, code, alpha_bl, alpha_t1, alpha_t2) %>%
  gather(key = measure, value = mean_score, - group, -code) %>%
  separate(measure, c("measure", "timepoint")) %>%
  mutate_if(is.character,funs(factor(.)))

hist(alpha_diff$mean_score)  

```


Design Mixed Linear Model (alpha power = dependent variable) 
```{r  Mixed Linear Model and Output}
  
  lmer_alpha<- lmer(alpha ~ timepoint*group + (1|code), 
              data = power_df_all) # timepoint fixed, group fixed, participant random                                                                    # timepoint*group interaction fixed, id = random
  summary(lmer_alpha) # output summary of model

  lmer_alpha_log<- lmer(log(alpha) ~ timepoint*group + (1|code), 
              data = power_df_all) # timepoint fixed, group fixed, participant random                                                                    # timepoint*group interaction fixed, id = random
  summary(lmer_alpha_log)
  
  model_ouptut<- tab_model(lmer_alpha, lmer_alpha_log, p.val = "kr", 
                           p.style = "both", 
                           show.df = TRUE, 
                           show.stat = TRUE)        # output in HTML format
                                                   # p-values is based on conditional F-tests with                                                             Kenward-Roger approximation for the degrees of                                                           freedom
  
            # for log(alpha) main effect of T1 and main effect of mtbi group?
  
  print(model_ouptut) #this won't work in markdown (html)
  
  # additional information  (optional)
  #confint(lmer_alpha) #calculate confidence intervals (2.5% is lower and 97.5% is upper bound)
  #coef(lmer_alpha_log) # summary of coefficients for each participants
  
  Anova(lmer_alpha_log, test="F")
  Anova(lmer_alpha, test="F") # all are significant (but non log transformed variables)
```

Check model assumptions/ run diagnositics
```{r Check assumptions}

plot(lmer_alpha_log)

diagnostics<-plot_model(lmer_alpha_log, type = "diag")

print(diagnostics)

# 4. Homogeneity of Variance (by statistical testing)
    # Regression models assume that variance of the residuals is equal across groups.
power_df_all$model.res<- residuals(lmer_alpha_log) #extracts the residuals and places them in a new column in our original data table
power_df_all$model.res.abs <-abs(power_df_all$model.res) #creates a new column with the absolute value of the residuals
power_df_all$model.res.2 <- power_df_all$model.res.abs^2 #squares the absolute values of the residuals to provide the more robust estimate
Levene.Model <- lm(model.res.2 ~ group, data=power_df_all) #ANOVA of the squared residuals
anova(Levene.Model) #displays the results
 #p value is 0.06 (so almost significant)

boxplot(resid(lmer_alpha_log)) 
```
  # assumptions all hold for alpha power.

```{r Exponentiate coefficients from log scale}

print(model_ouptut) # have a look at output ( # group is significant in this output- but not in below)

# need to exponentiate coefficients for interpretation purposes
# get info on main effect of group
(exp(coef(lmer_alpha_log)$code["groupmtbi"])- 1) * 100 # 66% increased in alpha power for a 1 unit                                                         increase in group ( e.g mTBI)

(exp(-0.01)- 1) * 100 # lower CI 
(exp(1.01)- 1) * 100 # upper CI 


# get info on interaction effect of group and timepoint
(exp(coef(lmer_alpha_log)$code["timepointT1"])- 1) * 100  # 22% decreased in alpha power at T1 (for both groups???)


(exp(-0.44)- 1) * 100 # lower CI 
(exp(-0.06)- 1) * 100 # upper CI 

(exp(coef(lmer_alpha_log)$code["timepointT1:groupmtbi"])- 1) * 100  # 24% decrease in alpha power at T1 for interaction 

(exp(-0.55)- 1) * 100 # lower CI 
(exp(0.001)- 1) * 100 # upper CI 

```  

```{r evaluate model and post hoc testing }

# main effects (no interaction effects)

library(phia)
lmer_alpha_means <- interactionMeans(lmer_alpha_log)
plot(lmer_alpha_means, traces=c("group","timepoint")) # create a plot (just for visualising)

```
Hmmmm.. so it looks as though there is a main effect of timepoint (ONLY) - with T1 alpha power being sig 
reduced compared to BL ( but no interaction- suggests its not to do with group).

When I do not log transform alpha, I see a significant effect of timepoint and timepoint*group- ...but then assumptions are all violated. Question for stats consult. 

```{r post hoc comparisons}
library(emmeans)

Anova(lmer_alpha_log, test= "F") # significant main effect of timepoint
#emeans_alpha<-emmeans(lmer_alpha_log, specs = pairwise ~ timepoint) # comparisons by timepoint

emeans_alpha<-emmeans(lmer_alpha_log, specs = pairwise ~ group|timepoint) # comparisons ( so can see only 
#These results are all on the model scale, so in this case these are estimated mean log response for each timepoint and group combination. 

emeans_alpha$emmeans # get means  for each comparison and CI

emeans_mfi$contrasts # The second part of the output, called contrasts, contains the comparisons of interest.


emeans_back_win<- emmeans(lmer_alpha_log, 
                      specs = pairwise ~ timepoint|group, 
                      type = "response") # timepoint comparisons

emeans_back_btw<- emmeans(lmer_alpha_log, 
                      specs = pairwise ~ group|timepoint, 
                      type = "response") # timepoint comparisons

#create a dataframe

alpha_contrasts <-emeans_alpha$contrasts %>%
     summary(infer=TRUE) %>%
     as.data.frame() %>%  #  p values are corrected for multiple comparisons but are in long form
     mutate(p.value = round(p.value, digits =4)) 


emeans_back_win$contrasts %>%
     rbind() %>% # bonferroni corrected- but for 3 tests ( good to use these)
     as.data.frame() %>%  #  p values are corrected for multiple comparisons but are in long form
     mutate(p.value = round(p.value, digits =4)) 

emeans_back_btw$contrasts %>%
     rbind() %>% # bonferroni corrected- but for 3 tests ( good to use these)
     as.data.frame() %>%  #  p values are corrected for multiple comparisons but are in long form
     mutate(p.value = round(p.value, digits =4)) 
        
```
  
  Optional additional analyses/visualisations
  
```{r Visualise marginal effects of model}

  g1<- plot_model(lmer_alpha_log, type = "pred", mdrt.values = "meansd") # can change type, "int" automatically plots                                                                          # interaction effects

  print(g1)

  p <- ggpredict(lmer_alpha_log, c("group", "timepoint"), type= "fe") # another way of visualising same info as above
  g2<- plot(p, add.data=FALSE)  # graph predicted values of alpha, change add data to TRUE to see data points
  
  print(g2)
  
  #The default, type = "fe", means that predictions are on the population-level and do not account 
  #for the random effect variances. Intervals are confidence intervals for the predicted values.

  #The idea behind the ggpredict function is to compute (and visualize) the relationship between a model predictor
  #(independent variable) and the model response (dependent variable).
  
  #Link below is info on customising plots and table output
 # https://strengejacke.github.io/sjPlot/articles/custplot.html
  
          
```
 # Resource on Model interpretation
 https://cran.r-project.org/web/packages/margins/vignettes/TechnicalDetails.pdf
 
 
```{r Visualising coefficients, include= FALSE}
 coeff_g<- dwplot(lmer_alpha_log)
  
  print(coeff_g)
  
  by_group<- power_df_all %>% 
    group_by(group) %>%                     # group data by group
    do(tidy(lm(alpha_log ~ timepoint, data = .))) %>% # run model on each grp
    dplyr::rename(model=group)  # make model variable
  
  coeff_by_grp<- dwplot(by_group, 
         vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) + # plot line at zero _behind_ coefs
    theme_bw() + xlab("Coefficient Estimate") + ylab("") +
    ggtitle("Predicting Alpha Power by Group") +
    theme(plot.title = element_text(face="bold"),
          legend.position = c(0.007, 0.01),
          legend.justification = c(0, 0),
          legend.background = element_rect(colour="grey80"),
          legend.title.align = .5) 
  
  print(coeff_by_grp)
  
```

Coefficients by fixed effects and interaction effects and then by group and timepoint

Visualise marginal effects
For MLM with interaction terms, marginal effects are considered easier to interpret than coefficients
See #https://cran.r-project.org/web/packages/ggeffects/vignettes/introduction_randomeffects.html
    #https://strengejacke.github.io/sjPlot/articles/plot_interactions.html for vignetttes.

```{r graph estimated marginal means }
alpha_graph<-emmip(lmer_alpha_log, group ~ timepoint, type = "response") + 
  aes(linetype = group, shape = group) + 
  scale_color_grey(start=0, end=0) +
  theme_bw() +
  labs(y = "Average Alpha Power",
       x = " ") +
   theme(
    legend.position = "bottom",
    legend.box.just = "center",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_blank()
    )

```

 