---
title: "Alpha_power_MLM"
author: "Hannah Coyle"
date: "18/10/2019"
output: html_document
---

```{r global options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figures/',
                      echo=FALSE, warning=FALSE, message=FALSE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
```

```{r load libaries, include = FALSE}

library(tidyverse)
library(knitr)
library(tidyverse)
library(Rmisc)
library(lme4)
library(sjPlot)
library(dotwhisker)
library(broom)
library(car)
library(e1071)
library(ggeffects)
library(magrittr)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(lme4)
```


# Load data frame with mean alpha power from ROI electrodes across BL, T1, T2 for controls and mTBI
```{r set up file}
here<-"~/Documents/PHD-Data-Analysis/PHD-Data-Analysis"
setwd(here) #main path/homedirectory
setwd("./Analysis/Data")
load("COMBINED_COG_PhD.Rdata")
#exclude participants 
COMBINED_COG_PhD <-subset(COMBINED_COG_PhD,!code %in% c(7, 16))
```


# Visualise data as bar and line graphs
```{r Visualise ravlt t1 perfomarnace longitudinally}

ravlt_df<-COMBINED_COG_PhD %>%
  dplyr::select(code, group, ravlt_t1_bl, ravlt_t1_t1, ravlt_t1_t2, wtar) 


ravlt_long <- ravlt_df %>%
  gather(key = measure, value = mean_score, - group, -code, - wtar) %>%
  separate(measure, c("measure", "domain", "timepoint")) %>%
  mutate_if(is.character,funs(factor(.))) %>%
  dplyr::select(-domain) 
  
ravlt_summary <- ravlt_long %>%
  group_by(group, timepoint) %>%
  summarySE(measurevar="mean_score", groupvars=c("group", "measure", "timepoint"), na.rm= TRUE)


# quick graph as a bar graph
ravlt_bar <- ravlt_summary %>%
  mutate(measure=factor(measure, labels ="ravlt")) %>%
  ggplot(aes(x=measure, y=mean_score, fill=group)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=mean_score-se, ymax=mean_score+se),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9))+
  ylab("Total words recalled") +
  facet_grid(~timepoint)+
  theme_light() +
  scale_fill_manual(values=c("deepskyblue1","royalblue4"))+
  ggtitle("Mean RAVLT score (trial 1)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title.x=element_blank(),
        axis.text.x=element_blank())

print(ravlt_bar)

 # quick graph as a line graph
  
  gbase<- ravlt_summary %>%
    ggplot(aes(y= mean_score, x= timepoint, colour=group))+
    geom_point()  #+ facet_grid(~group)
    #geom_errorbar(aes(ymin = alpha_power - se, ymax = alpha_power + se),
                  #width=.2) ## optional SE bars, make graph look messy though
  
  
  gline <- gbase + geom_line() 
  ravlt_summary$time = as.numeric(ravlt_summary$timepoint)
  gline <- gline %+% ravlt_summary
  
  print(gline + aes(x=time)+
          scale_x_continuous(breaks=c(1:3), labels=c("BL", "T1", "T2")))
  
```
 Bar graph and line graphs show distribution of means  
 
```{r}
## plot individual points with means as a bar plot
base_plot <- ggplot(ravlt_long, aes(x = timepoint, y = mean_score)) +
  facet_grid(~group) +
  stat_summary(aes(fill = factor(timepoint)), fun.y = mean, geom = "bar", color = "black", na.rm = TRUE) + 
  scale_fill_manual(values = c("#66c2a5", "#8da0cb", "#7f66c2")) +  # colorbrewer2.org
  theme(legend.position = "none")
# put individual lines on bar graph
subj_means_plot <- base_plot + 
  stat_summary(aes(group = code), fun.y = mean,  # means from raw data
               geom = "point", shape = 19, size = 4, color = "skyblue1") +
  stat_summary(aes(group = code), fun.y = mean, 
               geom = "line", size = 1.2, color = "skyblue1")

print(subj_means_plot)
```
 Bar graph with individual points shows individual trajectories (and what are likely to be outliers)


```{r check normality of outcome measre}
hist(ravlt_long$mean_score) # a little bit left skewed
hist(ravlt_long$wtar) # a little bit right skewed

#Try log transforming the outcome variable
ravlt_long$mean_score_log <-log(ravlt_long$mean_score)
ravlt_long$wtar_log <-log(ravlt_long$wtar)

hist(ravlt_long$mean_score_log) # more normally distributed- but with gaps
hist(ravlt_long$wtar_log) # makes it worse- so is it better to leave as is.

```

# Design Mixed Linear Model (ravlt_total_recalled = dependent variable) and visualise coefficient output
```{r  Mixed Linear Model and Output}
ravlt_long$code <-factor(ravlt_long$code)

lmer_ravlt<- lmer(mean_score ~ timepoint*group + wtar + (1|code), 
              data = ravlt_long, na.action=na.exclude) # timepoint fixed, group fixed, participant random                                                           # timepoint*group interaction fixed, id = random

#By default, lme4 assumes that all coefficients associated with the same random-effects term
#are correlated

lmer_ravlt_log<- lmer(log(mean_score) ~ timepoint*group + wtar + (1|code), 
              data = ravlt_long, na.action=na.exclude) #REML = FALSE (if want ML esti) 



lme_ravlt_log_covar<-lme(log(mean_score) ~ timepoint*group + wtar,random= ~1|code, data = ravlt_long, na.action=na.exclude, correlation= corCAR1(), method = "ML") 
                  # this is specifying the covariance structure, but is using nlme package

summary(lme_ravlt_log) # output summary of model
  
  model_ouptut<- tab_model(lmer_ravlt_log, lme_ravlt_log_covar,p.val = "kr", 
                           p.style = "both", show.stat =TRUE, 
                           show.df = TRUE)        # output in HTML format
 
  anova(lme_ravlt_log, lme_ravlt_log_1)                                                            
  
  # p-values is based on conditional F-tests with                                                                       Kenward-Roger approximation for the degrees of                                                                        freedom
  print(model_ouptut) #this won't work in markdown (html)
  
  
 
  # additional information  (optional)
  #confint(lmer_alpha) #calculate confidence intervals (2.5% is lower and 97.5% is upper bound)
  #coef(lmer_alpha) # summary of coefficients for each participants
  
  
```

For model output - will be in a seperate tab
The marginal R-squared considers only the variance of the fixed effects, while the conditional R-squared
takes both the fixed and random effects into account.
  
  

## Visualise marginal effects
For MLM with interaction terms, marginal effects are considered easier to interpret than coefficients
See #https://cran.r-project.org/web/packages/ggeffects/vignettes/introduction_randomeffects.html
    #https://strengejacke.github.io/sjPlot/articles/plot_interactions.html for vignetttes.
  
  
```{r Visualise marginal effects of model}

  g1<- plot_model(lmer_ravlt, type = "pred", mdrt.values = "meansd") # can change type, "int" automatically plots                                                                          # interaction effects

  print(g1)
  
  #another way of visualising same info

  p <- ggpredict(lmer_ravlt, c("group", "timepoint"), type= "fe") # another way of visualising same info as above
  g2<- plot(p, add.data=FALSE)  # graph predicted values of alpha, change add data to TRUE to see data points
  
  print(g2)
  
  #The default, type = "fe", means that predictions are on the population-level and do not account 
  #for the random effect variances. Intervals are confidence intervals for the predicted values.

  #The idea behind the ggpredict function is to compute (and visualize) the relationship between a model predictor
  #(independent variable) and the model response (dependent variable).
  
  #Link below is info on customising plots and table output
 # https://strengejacke.github.io/sjPlot/articles/custplot.html
  
  
          
```
 # Model interpretation
 https://cran.r-project.org/web/packages/margins/vignettes/TechnicalDetails.pdf
 
 What is the effect of timepoint and group of on ravlt performance?
  
  
  # Check model assumptions/ run diagnositics
```{r Check assumptions}

plot(lmer_ravlt_log) # normality of residuals

diagnostics<-plot_model(lmer_ravlt, type = "diag")

print(diagnostics)

#Homogeneity of Variance (by statistical testing)
    # Regression models assume that variance of the residuals is equal across groups.
ravlt_long$model.res<- residuals(lmer_ravlt) #extracts the residuals and places them in a new column in our original data table
ravlt_long$model.res.abs <-abs(ravlt_long$model.res) #creates a new column with the absolute value of the residuals
ravlt_long$model.res.2 <- ravlt_long$model.res.abs^2 #squares the absolute values of the residuals to provide the more robust estimate
Levene.Model <- lm(model.res.2 ~ group, data=ravlt_long) #ANOVA of the squared residuals
anova(Levene.Model) #displays the results
            #  assumption of homoscedasticity not violated !!!
# the p value is greater than 0.05, we can say that the variance of the residuals is equal and therefore the assumption of homoscedasticity is met - but is close p = 0.07


```


# B. Assess model fit
```{r Assess model fit}
# 5. Assess goodness-of-fit measures: R squared
cor(ravlt_long$mean_score_log, fitted(lmer(log(mean_score) ~ timepoint*group + (1|code),
                                  data = ravlt_long, na.action = na.exclude)), use= "complete.obs")^2

# 6. Assess data likelihood: What is the probability that we would
#observe the data we have given the model (i.e. given the
                                          #predictors we chose and given the ‘best’ parameter
                                          #estimates for those predictors).
summary(lmer_ravlt_log)
logLik(lmer_ravlt_log) # should always be negative, but closer to 0 is better
AIC(lmer_ravlt_log) # smaller is better (same with BIC)
```


```{r Exponentiate coefficients from log scale}

coef_st = tidy(lmer_ravlt, effects = "fixed",
     conf.int = TRUE,
     conf.method = "profile") # will only work if haven't run with lmerTest (creates a 6th column?)
# need to exponentiate coefficients for interpretation purposes

# get info on fixed effect of group
(exp(coef(lmer_ravlt_log)$code["groupmtbi"])- 1) * 100 # 14% decrease in ravlt t1 total score for a 1 unit                                                         increase in group ( e.g mTBI)

(exp(-0.29)- 1) * 100 # lower CI 
(exp(-0.01)- 1) * 100 # upper CI 

# fixed effect of group on html output
# no fixed effect of timepoint or interaction effect.



```

```{r model interpretation}
#Using Anova from car, we get p-values for the main effects.
Anova(lmer_ravlt_log, test = "F") # main effect of group not replicated 
                    #  The function Anova from car produces tables with p-values
                    # based on Wald tests, 
library(lmerTest) # will have to re run model 

#post hoc testing with emmeans (though potentially not relevant as no main effect of group or tp)
emeans_ravlt<-emmeans(lmer_ravlt_log,
                      specs = pairwise ~ group|timepoint)  #change | to : to look at all comparisons

#These results are all on the model scale, so in this case these are estimated mean log response 

emeans_ravlt$emmeans # get means  for each comparison and CI

emeans_ravlt$contrasts # The second part of the output, called contrasts, contains the comparisons of interest.

emeans_ravlt_back_trans<- emmeans(lmer_ravlt_log, specs = pairwise ~ group|timepoint, type = "response")

ravlt_contrasts <-emeans_ravlt$contrasts %>%
     rbind() %>% # bonferroni corrected- but for 3 tests ( good to use these)
     as.data.frame() %>%  #  p values are corrected for multiple comparisons but are in long form
     mutate(p.value = round(p.value, digits =4)) 
```

- group effect of mTBI - but only for some p value computations- also model does not seem to fit data very well, only explaining 44.7% of the variance.Log transformation only seems to improve the data a bit..? See a similar pattern of results for log transformed and non for ravlt_t1. Not sure how to report this finding? SHould be a simple main effect of group- but without means always a bit tricky to say where differene lies

```{r see if can improve model}
library(pbkrtest)
(fmLarge <- lmer((log(mean_score)) ~ timepoint*group + (1|code) + wtar, data = ravlt_long, na.action = na.exclude))


## removing Days
(fmSmall <- lmer((log(mean_score)) ~ timepoint+ (1|code) + wtar, data = ravlt_long, na.action = na.exclude))


anova(fmLarge,fmSmall) # not sig difference
n<-KRmodcomp(fmLarge,fmSmall)

getKR(n)
```


```{r graph estimated means from model}
ravlt_graph<-emmip(lmer_ravlt_log, group ~ timepoint, type = "response") + 
  aes(linetype = group, shape = group) + 
  scale_color_grey(start=0, end=0) +
  theme_bw() +
  labs(y = "RAVLT T1 total",
       x = " ") +
   theme(
    legend.position = "bottom",
    legend.box.just = "center",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_blank()
    )
```

  