---
title: "Demographics"
author: "Hannah Coyle"
email: "hannah.coyle@monash.edu"
date: "`r Sys.Date()`"
output: html_document
---
# Analysising clinical and cognitive data across three time points in a sample of mild traumatic brain injury (mtbi) patient.s

# Hannah Coyle, , January 2019
  #This r markdown document and output is for my demographic information. Includes summaries, tables etc

```{r global_option, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE, eval=TRUE)
```

# 1.Run normality tests on demographic information (to do)

# 2.Create demographics table for both groups
```{r Demographics table, include=FALSE}
library(tableone)
#change categorical variables to factors
myVars <- c("sex", "age", "education", "wtar")
catVars <- c("sex")

#create table by group
tab1 <- CreateTableOne(vars = myVars, strata = "group" , data = COMBINED_COG_PhD, factorVars = catVars)
print(tab1)

#if you want to specify nonnormal data you can do that as well (quote arguments avoids the mishandling of spaces)
  #create a variables that are nonnormal (BUT haven't checked that these aren't yet)
print(tab1, nonnormal= c("age", "education"), exact= c("wtar"), cramVars = "sex", quote= TRUE)

#view categorical and continuous parts of the table
tab1$CatTable
tab1$ContTable

#export into .csv/ excel and save table 
library(xlsx)
tab1Mat<- print(tab1, exact="stage", quote=TRUE, nospaces=TRUE,printToggle = FALSE)
write.csv(tab1Mat, file="demographics.csv")

```
#3.Create summary table for clinical characteristics (mtbi sample)